Verifying Spatial Properties of Array Computations

### 2017-04-16 oopsla17/study.tex
hypothesis 4 and 5 results

\begin{tabular}{l|l|l}
  (\textsf{const}) & Consistent & \fortran{a(i, j) = b(i, j) +
                                  b(i+1,j+1)} \\
  (\textsf{perm} & Permutation & \fortran{a(i, j) = c(j, i)} or \fortran{a(i,
                                                               0) =
                                 b(0, i)} \\
  (\textsf{LHSsub} & LHS subset & \fortran{a(i) = b(i, j) + b(i, j-1)}
  \\
  (\textsf{LHSsup} & LHS superset & \fortran{a(i, j) = b(i)} \\
  (\textsf{inconst} & Inconsistent & \fortran{a(i) = b(j)}
\end{tabular} \\[1em]
%

@@

\begin{tabular}{lS[table-format=4]@{\extracolsep{5pt}}S[table-format=2.2]c}

@@

Total                           & 108773    &  \\

@@

\begin{tabular}{lS[table-format=4]S[table-format=2.2]lS[table-format=4]S[table-format=2.2]}

@@

 \hline

@@

\begin{enumerate}[resume]
\item Many array computations are \emph{stencil computations}:
an assignment whose
 right-hand side comprises array subscripts with neighbourhood indices
and whose left-hand side is an neighbourhood-indexed array subscript
(\eg{}, the classic five-point stencil);
\end{enumerate}

>>>>

\begin{tabular}{l|l|l}
  (\textsf{const}) & Consistent & \fortran{a(i, j) = b(i, j) +
                                  b(i+1,j+1)} \\
  (\textsf{perm}) & Permutation & \fortran{a(i, j) = c(j, i)} or \fortran{a(i,
                                                               0) =
                                 b(0, i)} \\
  (\textsf{LHSsub}) & LHS subset & \fortran{a(i) = b(i, j) + b(i, j-1)}
  \\
  (\textsf{LHSsup}) & LHS superset & \fortran{a(i, j) = b(i)} \\
  (\textsf{inconst}) & Inconsistent & \fortran{a(i) = b(j)}
\end{tabular} \\[1em]
%

@@

\begin{tabular}{l|S[table-format=4]@{\extracolsep{5pt}}S[table-format=2.2]c}

@@

\textbf{Total}                  & 108773    &  \\

@@

\begin{tabular}{l|S[table-format=4]S[table-format=2.2]lS[table-format=4]S[table-format=2.2]}

@@

 \hline \hline

@@

\item Many array computations are \emph{stencil computations}:
an assignment whose
 right-hand side comprises array subscripts with neighbourhood indices
and whose left-hand side is an neighbourhood-indexed array subscript
(\eg{}, the classic five-point stencil);
We subclassified the data based on the right-hand side
and left-hand side categories
being either a variable, or induction-variable indexed,
neighbour-indexed or affined-indexed array. We additionally
included an orthogonal classification of whether
left-hand and right-hand sides are consistent (have a common subset
of induction variables) or not.
\begin{tabular}{ll|S[table-format=4]S[table-format=2.2]S[table-format=4]S[table-format=2.2]c}
\multicolumn{2}{l|}{\textbf{LHS and RHS classification}}
& \textbf{\# inconsistent}
& \% {(whole)}
& \textbf{\# consistent}
& \% {(whole)}
& \textit{stencils} \\ \hline
LHS Vars & RHS neigh & 0 & 0.00\% & 4591      & 4.22\%
 & \multirow{7}{*}{66.01\%} \\\cline{1-6}
LHS IVs & RHS neigh   &  4560      & 4.19\%  & 50861     & 46.76\% \\\cline{1-6}
LHS neigh & RHS neigh & 2358      & 2.17\% & 16336     & 15.02\%   \\\cline{1-6}
LHS Vars & RHS affines & 0 & 0.00\%  & 4         & 0.00\% \\\cline{1-6}
LHS IVs, & RHS affines & 7 & 0.01\% & 0 & 0.00\%  \\\cline{1-6}
LHS affine & RHS affines & 0 & 0.00\% &  6         & 0.01\% \\\hline
LHS neigh & RHS affines & 8         & 0.01\% & 0 & 0.00\%  \\\hline
LHS affine & RHS neigh  & 23        & 0.02\% & 0 & 0.00\% \\\hline \hline
other       &                      & 30019     & 27.60\%  & &  \\\hline%\hline
%Total        &                      & 108773    &
%  LHS var, & RHS affine             & 4         & 0.00\%
% & \multirow{7}{*}{72.37\%} \\\cline{1-3}
%  LHS var, & RHS neighbour          & 4591      & 4.22\%  \\\cline{1-3}
%  LHS IVs, &RHS affine             & 7         & 0.01\%  \\\cline{1-3}
%  LHS IVs, &RHS neighbour          & 55421     & 50.95\% \\\cline{1-3}
%  LHS neighbour, &RHS neighbour    & 18694     & 17.19\% \\\cline{1-3}
%  LHS affine, &RHS affine          & 6         & 0.01\%  \\\hline
%  LHS affine, &RHS neighbour       & 23        & 0.02\%  \\\hline
%  LHS neighbour, &RHS affine       & 8         & 0.01\%  \\\hline
%  other   & other                   & 30019     & 27.60\%  %\\\hline\hline
%  Total                           & 108773    &  \\
The most common category has an l.h.s array subscript
indexed by induction variables, and a right-hand side comprising
neighbourhood offsets. Next most common is to have neighbourhood
offsets on the left-hand and right-hand sides. Notably
there are two categories always seen as inconsistent:
``LHS affine, RHS neighbour'' and ``LHS neighbour, RHS affine''
since they comprise different indexing schemes
on each side of the assignment, e.g. $a(i + 1) = b(2*i + 1)$. These
appear very rarely. Thus, we find that $66.01\%$ of array computations
have a static pattern on the left and right-hand side that is either
affine or a constant translation. Thus, the data support ours
hypothesis that stencis are common. Based on this grouping, we
considered the two subclassifications of different kinds of
left-right-hand side relationship and contiguity:
\begin{minipage}{1\linewidth}
\begin{tabular}{l|S[table-format=4]S[table-format=2.2]}
\textbf{Contiguity} & \textbf{Number} & \% {(whole)} \\\hline
Contiguous stencil         & 68246     & 62.74\%  \\\hline
Non-contiguous stencil     & 3552      & 3.27\%  \\\hline
other                      & 36975     & 33.99\% \\\hline
%\textbf{Total}             & 108773    &  \\
\hspace{1em}
\begin{tabular}{l|S[table-format=4]S[table-format=2.2]S[table-format=4]S[table-format=2.2]}
\textbf{Consistency} & \textbf{Number} & \% & \textbf{relativised \#} & \% \\\hline
Direct                            & 47448     & 43.62\% & 636       & 0.58\%  \\\hline
LHSsubset                         & 9812      & 9.02\% & 65        & 0.06\%  \\\hline
LHSsuperset                       & 10110     & 9.29\%  & 84        & 0.08\%  \\\hline
Permutation                       & 3604      & 3.31\%  & 39        & 0.04\%  \\\hline
other                             & 36975     & 33.99\% \\\hline
\end{minipage}
Most stencils are contiguous, but there is a non trivial
amount of non-contiguity. Most stencils have matching induction
variables on the left and right-hand sides but again there
are non-trivial amounts of other kinds of relationship.
\item Many array computations read from each particular index just
  once.
We classified right-hand sides which are neighbourhood offsets (or
affine) and subclassified this by whether array subscripts were
unique or repeated:

\begin{center}
\begin{tabular}{l|S[table-format=4]S[table-format=2.2]S[table-format=4]}
\textbf{RHS form and reuse} & \textbf{Number} & \% \\ \hline
RHS neighbour/affine unique indices      & 78349     & 72.03\%  \\\hline
RHS neighbour/affine repeated indices   & 10228     & 9.40\%  \\\hline
other                             & 20196     & 18.57\%  \\\hline\hline
Total                             & 108773    &
\end{tabular}
\end{center}
%
Thus, hypothesis 5 is confirmed, the majority of array computations
have access patterns with linear usage of array subscripts.
